---
title: "20210629 - Rescue Animals"
author: "Nick Cruickshank"
date: "6/30/2021"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

```{r}
# libraries
library(forcats)
library(lubridate)
library(readr)
library(tidyverse)
```

```{r}
# data
animal_rescues <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-06-29/animal_rescues.csv')
```

# Introduction

Data this week comes from [London.gov](https://data.london.gov.uk/dataset/animal-rescue-incidents-attended-by-lfb). 

> Fox in bedroom, dog trapped in wall. The London Fire Brigade responds to hundreds of requests to rescue animals each year. Its monthly-updated spreadsheet of such events goes back to 2009; it lists the location and type of property, the kind of animal and rescue, hours spent, a (very) brief description, and more. [h/t Soph Warnes]

Another [article](https://www.theguardian.com/world/2021/jan/08/animal-rescues-london-fire-brigade-rise-2020-pandemic-year) found that animals rescues increased by 20% in the year of the pandemic (2020). 

> The London fire brigade (LFB) was involved in 755 such incidents – more than two a day. The number of rescues rose by 20% compared with 2019 when there were 602, with the biggest rise coming in the number of non-domestic animals rescued, according to the data.

```{r}
# tidy
animal_rescues

unredacted <- animal_rescues %>%
  filter(final_description != "Redacted")
```

# Exploratory Data Analysis

## Notes

```{r}
hist(animal_rescues$cal_year)
```

`type_of_incident`: Only 'Special Service' represented.

`pump` = truck / unit deployed.

```{r fig.height=4, fig.width=10}
animal_rescues %>%
  count(originof_call) %>%
  ggplot(aes(fct_reorder(originof_call, -n), n)) + 
  geom_bar(stat = "identity")
```


# Data Visualization Projects

## Choropleth of animal rescues

## Timeline of animal rescue

By species?