---
title: "20210622 - Parks Access"
author: "Nick Cruickshank"
date: "6/23/2021"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

```{r}
# library
library(cowplot)
library(forcats)
library(glue)
library(readr)
library(tidyverse)
```

```{r}
# data
parks <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-06-22/parks.csv')
```

# Introduction

This weeks project uses data sourced from [The Trust for Public Land](https://www.tpl.org/parks-and-an-equitable-recovery-parkscore-report). The article written by [Citylab](https://www.bloomberg.com/news/articles/2021-05-27/the-cities-where-people-of-color-can-walk-to-a-park) provides a good starting narrative for park access and this dataset.

## Noteworthy Excerpts

> When the pandemic forced millions of Americans to shelter indoors, parks saw a surge in popularity. Parks in North Carolina received an all-time high of 20 million visitors in 2020, an uptick of one million from the year before. In New York, the number of park-goers topped 78 million. 

> But even as green spaces proved to be a crucial element to people’s physical and mental well-being — especially for urban dwellers — the pandemic further exposed the disparity in who has access to parks. ... Low-income communities have access to 42% less than high-income neighborhoods.

> The Trust for Public Land (TPL) has kept track of green space availability across U.S. metros, using the ParkScore index, which measures how well cities are meeting their residents’ need for parks based on [five] metrics: park access, acreage, investment, amenities, ... [and] equity.

> Its assessment is grounded in the idea that Americans should have park access within a 10-minute walk from their home. 

# Exploratory Analysis

This data set explores an `r max(parks$year) - min(parks$year)` timeframe from `r min(parks$date)` to `r max(parks$date)`.

## Tidy Data Set

```{r}
# Tidy
tidy_parks <- parks %>%
  mutate(
    park_pct_city_data = as.double(str_remove(park_pct_city_data, "\\%")),
    pct_near_park_data = as.double(str_remove(pct_near_park_data, "\\%"))
    )

city_avg_points <- parks %>%
  group_by(city) %>%
  dplyr::summarise(
    years = n(),
    avg_total_points = mean(total_points, na.rm = TRUE)
    ) %>%
  filter(years > 2) %>%
  arrange(desc(avg_total_points))

top_25_cities <- head(city_avg_points, 25)
top_10_cities <- head(city_avg_points, 10)
```

## Yearly rank of cities over time

Plot year by total points, facet by city. Which cities have gone up or down

```{r top_city_rank_over_time, fig.height=4, fig.width=10}
parks %>%
  filter(
    city %in% top_10_cities$city
  )  %>%
  left_join(city_avg_points) %>%
  ggplot(aes(year, total_points)) + 
  geom_line() + 
  geom_point() + 
  geom_hline(aes(yintercept = avg_total_points), linetype = "dotdash") +
  labs(
    title = "Total park points over time for the cities with the highest average total points",
    subtitle = "Dashed horizontal line represents average total points"
  ) +
  facet_wrap(~ city, ncol = 5) + 
  theme_bw()
```

Data points don't look particularly well distributed. Each of these cities seems to have surged in 2020. I wonder if that is an anomally of these cities, or a trend seen throughout the dataset.

```{r avg_total_rank_over_time, fig.height=4, fig.width=10}
parks %>%
  group_by(year) %>%
  dplyr::summarise(avg_total_points = mean(total_points, na.rm = TRUE)) %>%
  ggplot(aes(year, avg_total_points)) + 
  geom_smooth(se = FALSE) + 
  geom_line(size = 1.2) + 
  geom_point(size = 2) + 
  labs(
    title = "Average total points for all cities over time",
    subtitle = "Dataset supports the idea that total points assigned by the TPL during assessments grew a lot in 2020",
    caption = "Looks like the dataset does not include 2014 and 2019 data"
  ) + 
  theme_bw()
```


## Which cities have the largest parks?

Median park size

```{r top_cities_w_most_parks, fig.height=6, fig.width=12}
top_20_parks_med_size <- parks %>%
  filter(year == 2020) %>%
  arrange(desc(med_park_size_data)) %>%
  head(20) %>%
  ggplot(aes(fct_reorder(city, med_park_size_data), med_park_size_data)) + 
  geom_bar(stat = "identity", fill = "cadetblue2", alpha = 0.8, color = "cadetblue4", size = 1.05) + 
  geom_text(aes(label = paste0(med_park_size_data, " acres")), hjust = 1.1) +
  coord_flip() + 
  labs(
    title = "Top 20 Cities in 2020 by Median Park Size"
  ) + 
  theme_minimal() +
  theme(
    axis.title = element_blank(),
    panel.grid = element_blank(),
    axis.text.x = element_blank()
  )

top_20_parks_perc_size <- parks %>%
  filter(year == 2020) %>%
  mutate(park_pct_city_data = as.double(str_remove(park_pct_city_data, "\\%"))) %>%
  arrange(desc(park_pct_city_data)) %>%
  head(20) %>%
  ggplot(aes(fct_reorder(city, park_pct_city_data), park_pct_city_data)) + 
  geom_bar(stat = "identity", fill = "palegreen2", alpha = 0.8, color = "forestgreen", size = 1.05) + 
  geom_text(aes(label = paste0(park_pct_city_data, "%")), hjust = 1.1) +
  coord_flip() + 
  labs(
    title = "Top 20 Cities in 2020 by Park % City Size"
  ) + 
  theme_minimal() +
  theme(
    axis.title = element_blank(),
    panel.grid = element_blank(),
    axis.text.x = element_blank()
  )

park_size_plot <- plot_grid(top_20_parks_med_size, top_20_parks_perc_size) + 
  theme(
    plot.background = element_rect(fill = "khaki")
  )

park_size_plot
```


## Which cities have the highest percent of residents within a 10min walk?

### Distribution

```{r park_access_linerange, fig.height=6, fig.width=8}
access_label <- data.frame(city = "San Francisco", median = 42,
                           label = "Percent of city residents\nwho live 10 minutes away\nfrom a park (2012 - 2020)")

parks %>%
  filter(
    city %in% top_25_cities$city
  ) %>%
  mutate(pct_near_park_data = as.double(str_remove(pct_near_park_data, "\\%"))) %>%
  group_by(city) %>%
  dplyr::summarise(
    min = min(pct_near_park_data),
    median = median(pct_near_park_data),
    max = max(pct_near_park_data)
  ) %>%
  ggplot(aes(fct_reorder(city, median), median)) + 
  geom_pointrange(aes(ymin = min, ymax = max), color = "royalblue", size = 1.1) + 
  geom_text(aes(y = min - 1, label = city), hjust = "right", color = "burlywood4", fontface = "bold") + 
  geom_text(aes(y = max + 1, label = paste0(median, "%")), hjust = "left", color = "burlywood4", fontface = "bold") + 
  geom_text(data = access_label, aes(label = label), vjust = "top",
            size = 6, fontface = "bold", color = "forestgreen") +
  coord_flip(ylim = c(30, 102)) +
  theme_void() + 
  theme(
    plot.background = element_rect(fill = "palegreen1", color = "forestgreen", size = 2)
  )
```


### 10min Walk by Percent City Size

```{r park_access_by_city_park_area, fig.height=4, fig.width=8}
# make linear model
park_access_model1 <- lm(formula = park_pct_city_data ~ pct_near_park_data, data = tidy_parks)

# get formula
park_access_model1_intercept <- park_access_model1$coefficients["(Intercept)"]
park_access_model1_weight <- park_access_model1$coefficients["pct_near_park_data"]
park_access_model1_rsq <- summary(park_access_model1)$r.squared

tidy_parks %>%
  ggplot(aes(pct_near_park_data, park_pct_city_data)) + 
  geom_smooth(method = "lm", se = FALSE, color = "royalblue", size = 2) + 
  geom_point(color = "forestgreen", size = 2, alpha = 0.8) + 
  labs(
    title = "Surprisingly, cities residents have greater access to parks when cities have more parks",
    subtitle = glue("Parkland % City Area = {round(park_access_model1_intercept, 2)} + {round(park_access_model1_weight, 2)}[% Residents Near Park] (R Squared = {round(park_access_model1_rsq, 2)})"),
    x = "% Residents Near Park",
    y = "Parkland % City Area"
  ) + 
  theme_bw() + 
  theme(
    panel.background = element_rect(fill = "palegreen1"),
    plot.background = element_rect(fill = "khaki", color = "burlywood4", size = 2)
  )
```


The 45%+ `Parkland % City Area` might be outliers, and should be filtered out for this graph.

## Distribution of auxillary park types by city

Pivot `basketball_data`, `dogpark_data`, `playground_data`, `rec_sr_data`, and `splashground_data`.
First you'll need to standardize the per capita metrics.

## Relationship between spending per resident and total rank

# Primary Plotting Project