---
title: "20210727 - Olympics"
author: "Nick Cruickshank"
date: "7/31/2021"
output: 
  github_document:
    toc: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

```{r libraries}
# libraries
library(ggtext)
library(ggstream)
library(readr)
library(stringi)
library(tidyverse)
```

```{r data}
# data
olympics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-07-27/olympics.csv')

olympic_doping <- read_csv("olympic_doping.csv") %>%
  janitor::clean_names()
```

# Ideas for Visualization

- Proportion of Gold Medals Over Time
   - For each year of the olympics, find the proportion of medals for each of the Top X nations  
   - Include an other stream for not Top X nations  
- Russian Blood Doping
   - Do some kind of timeline analysis for Russian athletes  
   - Be sure to replace 'Soviet Union' with 'Russia'
- Distributions of Weights and/or Heights by Sport


# Gold Medal Proportions


# Russian Blood Doping

[Wikipedia Page for Blood Doping Scandal](https://en.wikipedia.org/wiki/Doping_in_Russia)

```{r}
rus <- olympics %>%
  mutate(team = ifelse(team == "Soviet Union", "Russia", team)) %>%
  filter(
    season == "Summer",
    team == "Russia"
  ) %>%
  group_by(year, city) %>% #and by sport?
  dplyr::summarise(
    athletes = n_distinct(name),
    gold = sum(medal == "Gold", na.rm = TRUE),
    silver = sum(medal == "Silver", na.rm = TRUE),
    bronze = sum(medal == "Bronze", na.rm = TRUE),
    non_gold = sum(medal == "Silver" | medal == "Bronze", na.rm = TRUE)
  )

olympic_game_segments <- tibble(x = c(seq(1896, 1912, by = 4), seq(1920, 1936, by = 4), seq(1948, 2016, by = 4)), y = 3)

## retrieve olympic cities and years
summer_og_editions <- olympics %>%
  filter(season == "Summer") %>%
  distinct(year, city) %>%
  mutate(edition = glue::glue("{city}\n{year}")) %>%
  filter(
    !(year == 1906 & city == "Athina"),
    !(year == 1956 & city == "Stockholm")
  ) %>%
  arrange(year) %>%
  rowid_to_column() 
```

```{r Russian Medals by Blood Doping, fig.height=7, fig.width=18}
label_segment <- tibble(x = c(1956, 2008),
                        xend = c(1970, 2018),
                        y = c(72, 52),
                        yend = c(72, 52))

label_arrows <- tibble(x = c(1970, 2016),
                       xend = c(1978, 2016),
                       y = c(72, 52),
                       yend = c(83, 43))

rus %>%
  pivot_longer(cols = c("gold", "non_gold")) %>% #c("gold", "silver", "bronze")
  mutate(
    ratio = 100 * (value / athletes),
    name = factor(name, levels = c("gold", "non_gold"))
    #name = factor(str_to_title(name), levels = c("Gold", "Silver", "Bronze"))
  ) %>%
  ggplot() + 
  # dashed x grid
  geom_segment(data = olympic_game_segments, aes(x = x, xend = x, y = 0, yend = y), size = 0.5, color = "white", linetype = "dashed") +
  # details x axis labels
  geom_text(data = summer_og_editions, aes(x = year, y = -7, label = edition), 
            angle = 45, lineheight = 0.8, fontface = "bold", color = "white") +
  geom_col(aes(year, ratio, fill = name), color = "grey20") + 
  scale_fill_manual(values = c(
    "gold" = "gold3",
    "non_gold" = "gray40"
  )) + 
  # seqments for pointing out interesting bars
  geom_segment(data = label_segment, aes(x = x, xend = xend, y = y, yend = yend), color = "white") + 
  geom_segment(data = label_arrows, aes(x = x, xend = xend, y = y, yend = yend), color = "white", arrow = arrow(length = unit(2, "mm"))) +
  # title
  annotate("text", x = 1918, y = 83, label = "Percent of Russian Athletes to Medal", size = 10, fontface = "bold", color = "white") + 
  annotate("text", x = 1918, y = 73, label = "How did the Russian Blood Doping Scandal\naffect performance?", lineheight = 0.95, size = 8, fontface = "italic", color = "white") +
  # subtitle
  # notes
  annotate(geom = "text", x = 1916, y = 10, label = "WORLD WAR I\n(1914-1918)", fontface = "bold", size = 5, color = "white") + 
  annotate(geom = "text", x = 1942, y = 10, label = "WORLD WAR II\n(1939-1945)", fontface = "bold", size = 5, color = "white") +
  annotate(geom = "text", x = 1960, y = 95, label = "The Chemists' Games", fontface = "bold", color = "white", size = 4.5) + 
  annotate(geom = "text", x = 1960, y = 83, label = "No athletes caught during the games.\nAdditional testing afterwards using new\ntechniques revealed 20% of specimens would\nhave resulted in disciplinary action.\nFirst documented case of blood doping.", color = "white") +
  annotate(geom = "text", x = 2012, y = 60, label = "38% of Russian atheletes submitted\nfor competition were rejected due\nto state-sponsored doping scandal", color = "white", size = 4.5) +
  labs(
    caption = "**Source:** sports-reference.com & Kaggle  |  **More Info:** en.wikipedia.org/wiki/Doping_at_the_Olympic_Games  |  **Visualization By:** Nick Cruickshank"
  ) +
  coord_cartesian(ylim = c(-10, 100)) +
  scale_x_continuous(name = NULL) +
  scale_y_continuous(name = NULL, labels = function(x) paste0(x, "%")) +
  theme(
    plot.background = element_rect(fill = "firebrick4"),
    panel.background = element_rect(fill = "firebrick4"),
    panel.grid = element_blank(),
    legend.position = "none",
    axis.text.y = element_text(color = "white", size = 15),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    plot.caption = element_markdown(color = "white", size = 12)
  )
```

- Embellish this plot with historical info at relevant bars   
   - Russia has had the most medals stripped of any country (13 medals)
- Replace x axis with "City\nYear"
- Highlight years where the most russian athletes won, label with the sport  

```{r}
dope <- olympic_doping %>%
  fill(c(name, country), .direction = "down") %>%
  filter(str_detect(country, "(Russia|Soviet Union)")) %>%
  mutate(
    medal = case_when(
      str_detect(medals, "gold") ~ "Gold",
      str_detect(medals, "silver") ~ "Silver",
      str_detect(medals, "bronze") ~ "Bronze"
    )
  ) %>%
  select(year, city, anti_doping_rule_violation, sport, dns, medal)

sort(unique(rus_ath_year$sport))

sort(unique(dope$sport))
```

```{r}
# bar chart of russian dopers at each games
## y axis in percent of total atheletes
## fill with sport
## inverted bar chart to be placed below the primary plot

## need to make the 'dope' sports equal to olympic sports

rus_ath_year <- olympics %>%
  filter(
    season == "Summer", 
    str_detect(team, "(Russia|Soviet Union)")
  ) %>%
  group_by(year) %>% # and by sport if possible
  dplyr::summarise(
    athletes = n_distinct(name)
  )

dope %>%
  count(year, city) %>% # and by sport if possible 
  inner_join(rus_ath_year) %>%
  mutate(prop = 100 * (n / athletes)) %>%
  ggplot(aes(year, -prop)) + 
  geom_col(fill = "dodgerblue") + 
  scale_y_continuous(name = NULL, labels = function(x) paste0(abs(x), "%")) + 
  coord_cartesian(xlim = c(1896,2016)) + 
  theme(
    plot.background = element_rect(fill = "firebrick4"),
    panel.background = element_rect(fill = "firebrick4"),
    panel.grid = element_blank(),
    axis.text.y = element_text(color = "white", size = 15),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
  )

```

```{r}
# pie chart of which drugs used by dopers
## will likely need to tidy 'anti_doping_rule_violation' column and separate_rows()
```

```{r}
# pie chart of which sports had the most dopers
```


# Top Medaling Winter Olympics

- Try grouping by event to count the numnber of events with medalss

```{r}
olympics %>%
  filter(season == "Winter") %>%
  mutate(
    team = case_when(
      team == "Soviet Union" ~ "Russia",
      team %in% c("East Germany", "West Germany") ~ "Germany",
      team == "United States-1" ~ "United States",
      team == "China-1" ~ "China",
      team == "Australasia" ~ "Australia",
      TRUE ~ team
    )
  ) %>%
  group_by(team) %>%
  dplyr::summarise(
    total_athletes = n(),
    total_gold = sum(medal == "Gold", na.rm = TRUE),
    total_silver = sum(medal == "Silver", na.rm = TRUE),
    total_bronze = sum(medal == "Bronze", na.rm = TRUE),
    total_medals = total_gold + total_silver + total_bronze
  ) %>%
  arrange(desc(total_medals))
```

