---
title: "20210727 - Olympics"
author: "Nick Cruickshank"
date: "7/31/2021"
output: 
  github_document:
    toc: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

```{r libraries}
# libraries
library(ggtext)
library(ggstream)
library(readr)
library(tidyverse)
```

```{r data}
# data
olympics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-07-27/olympics.csv')
```

# Ideas for Visualization

- Proportion of Gold Medals Over Time
   - For each year of the olympics, find the proportion of medals for each of the Top X nations  
   - Include an other stream for not Top X nations  
- Russian Blood Doping
   - Do some kind of timeline analysis for Russian athletes  
   - Be sure to replace 'Soviet Union' with 'Russia'
- Distributions of Weights and/or Heights by Sport


# Gold Medal Proportions


# Russian Blood Doping

[Wikipedia Page for Blood Doping Scandal](https://en.wikipedia.org/wiki/Doping_in_Russia)

```{r}
rus <- olympics %>%
  mutate(team = ifelse(team == "Soviet Union", "Russia", team)) %>%
  filter(
    season == "Summer",
    team == "Russia"
  ) %>%
  group_by(year, city) %>% #and by sport?
  dplyr::summarise(
    athletes = n(),
    gold = sum(medal == "Gold", na.rm = TRUE),
    silver = sum(medal == "Silver", na.rm = TRUE),
    bronze = sum(medal == "Bronze", na.rm = TRUE),
    non_gold = sum(medal == "Silver" | medal == "Bronze", na.rm = TRUE)
  )

olympic_game_segments <- tibble(x = c(seq(1896, 1912, by = 4), seq(1920, 1936, by = 4), seq(1948, 2016, by = 4)), y = 3)

## retrieve olympic cities and years
summer_og_editions <- olympics %>%
  filter(season == "Summer") %>%
  distinct(year, city) %>%
  mutate(edition = glue::glue("{city}\n{year}")) %>%
  filter(
    !(year == 1906 & city == "Athina"),
    !(year == 1956 & city == "Stockholm")
  ) %>%
  arrange(year) %>%
  rowid_to_column() 
```

```{r Russian Medals by Blood Doping, fig.height=7, fig.width=18}
# label_segment <- tibble(x = c(1969, 1973, 1984, 1969.5,1973.5, 1900),
#                         xend = c(1969,1973, 1990, 1974, 1976, 1912),
#                         y = c(160,200, 208, 140, 185, 75),
#                         yend = c(125,170, 208, 140, 185, 75))
# 
# label_arrows <- tibble(x = c(1974,1976, 1988, 1906, 1906),
#                        xend = c(1976,1980, 1993.5, 1906, 1906),
#                        y = c(140,185,205, 72, -55),
#                        yend = c(105,100, 185, 30, -25))

label_segment <- tibble(x = c(1956),
                        xend = c(1970),
                        y = c(50),
                        yend = c(50))

label_arrows <- tibble(x = c(1970),
                       xend = c(1976),
                       y = c(50),
                       yend = c(60))

rus %>%
  pivot_longer(cols = c("gold", "non_gold")) %>% #c("gold", "silver", "bronze")
  mutate(
    ratio = 100 * (value / athletes),
    name = factor(name, levels = c("gold", "non_gold"))
    #name = factor(str_to_title(name), levels = c("Gold", "Silver", "Bronze"))
  ) %>%
  ggplot() + 
  # dashed x grid
  geom_segment(data = olympic_game_segments, aes(x = x, xend = x, y = 0, yend = y), size = 0.5, color = "white", linetype = "dashed") +
  # details x axis labels
  geom_text(data = summer_og_editions, aes(x = year, y = -7, label = edition), 
            angle = 45, lineheight = 0.8, fontface = "bold", color = "white") +
  geom_col(aes(year, ratio, fill = name), color = "grey20") + 
  scale_fill_manual(values = c(
    "gold" = "gold3",
    "non_gold" = "gray40"
  )) + 
  # seqments for pointing out interesting bars
  geom_segment(data = label_segment, aes(x = x, xend = xend, y = y, yend = yend), color = "white") + 
  geom_segment(data = label_arrows, aes(x = x, xend = xend, y = y, yend = yend), color = "white", arrow = arrow(length = unit(2, "mm"))) +
  # title
  annotate("text", x = 1918, y = 65, label = "Percent of Russian Athletes to Medal", size = 10, fontface = "bold", color = "white") + 
  annotate("text", x = 1918, y = 57, label = "How did the Russian Blood Doping Scandal\naffect performance?", lineheight = 0.95, size = 8, fontface = "italic", color = "white") +
  # subtitle
  # notes
  annotate(geom = "text", x = 1916, y = 10, label = "WORLD WAR I\n(1914-1918)", fontface = "bold", size = 5, color = "white") + 
  annotate(geom = "text", x = 1942, y = 10, label = "WORLD WAR II\n(1939-1945)", fontface = "bold", size = 5, color = "white") +
  annotate(geom = "text", x = 1960, y = 67, label = "The Chemists' Games", fontface = "bold", color = "white", size = 4.5) + 
  annotate(geom = "text", x = 1960, y = 58, label = "No athletes caught during the games.\nAdditional testing afterwards using new\ntechniques revealed 20% of specimens would\nhave resulted in disciplinary action.\nFirst documented case of blood doping.", color = "white") +
  labs(
    caption = "**Source:** sports-reference.com & Kaggle  |  **More Info:** en.wikipedia.org/wiki/Doping_at_the_Olympic_Games  |  **Visualization By:** Nick Cruickshank"
  ) +
  coord_cartesian(ylim = c(-10, 70)) +
  scale_x_continuous(name = NULL) +
  scale_y_continuous(name = NULL, labels = function(x) paste0(x, "%")) +
  theme(
    plot.background = element_rect(fill = "firebrick4"),
    panel.background = element_rect(fill = "firebrick4"),
    panel.grid = element_blank(),
    legend.position = "none",
    axis.text.y = element_text(color = "white", size = 15),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    plot.caption = element_markdown(color = "white", size = 12)
  )
```

- Embellish this plot with historical info at relevant bars   
   - Originally, Russia submitted a list of 389 athletes for competition. On 7 August 2016, the IOC cleared 278 athletes, and 111 were removed because of the state-sponsored doping scandal.
- Replace x axis with "City\nYear"
- Highlight years where the most russian athletes won, label with the sport  

# Top Medaling Winter Olympics

- Try grouping by event to count the numnber of events with medalss

```{r}
olympics %>%
  filter(season == "Winter") %>%
  mutate(
    team = case_when(
      team == "Soviet Union" ~ "Russia",
      team %in% c("East Germany", "West Germany") ~ "Germany",
      team == "United States-1" ~ "United States",
      team == "China-1" ~ "China",
      team == "Australasia" ~ "Australia",
      TRUE ~ team
    )
  ) %>%
  group_by(team) %>%
  dplyr::summarise(
    total_athletes = n(),
    total_gold = sum(medal == "Gold", na.rm = TRUE),
    total_silver = sum(medal == "Silver", na.rm = TRUE),
    total_bronze = sum(medal == "Bronze", na.rm = TRUE),
    total_medals = total_gold + total_silver + total_bronze
  ) %>%
  arrange(desc(total_medals))
```

