---
title: "Tidy Tuesday 20210601 - SURVIVOR"
author: "Nick Cruickshank"
date: "6/1/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

```{r}
# libraries
library(cowplot)
library(forcats)
library(ggrepel)
library(readr)
library(tidytext)
library(tidyverse)
```

```{r}
# data
summary <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-06-01/summary.csv')
castaways <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-06-01/castaways.csv')
jury_votes <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-06-01/jury_votes.csv')
challenges <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-06-01/challenges.csv')
viewers <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-06-01/viewers.csv')
```

```{r}
# functions
Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}
```

# Exploratory Analysis

## Summary

```{r}
knitr::kable(head(summary),
             align = 'c',
             caption = "Head of summary table")
```

Which season had the greatest disparity between viewers at premier vs at finale?

```{r fig.height=5, fig.width=10}
# identify what the top five countries are for filling the graph
list_top5_countries <- head(arrange(count(summary, vars = country), desc(n)), 5)$vars

# reshape summary by country with "Other" category
sum_pie_df <- summary %>%
  mutate(
    country_category = ifelse(country %in% list_top5_countries, country, "Other")
  ) %>%
  group_by(country_category) %>%
  dplyr::summarise(value = n())

# colors to use by country
countries_colors <- c(
  "Brazil" = "limegreen",
  "Fiji" = "deepskyblue",
  "Nicaragua" = "royalblue",
  "Other" = "gray50",
  "Panama" = "red",
  "Philippines" = "gold"
)

# create plots
## pie chart and legen
sum_pie <- ggplot(sum_pie_df, aes(x = "", y = value, fill = country_category)) + 
  geom_bar(stat = "identity", color = "gray20") + 
  coord_polar("y", start = 0) + 
  scale_fill_manual(values = countries_colors) + 
  labs(title = "Location Breakdown", fill = "Country") +
  theme_void() + 
  theme(legend.position = "bottom")

## viewership delta
sum_delta_plot <- summary %>%
  mutate(
    viewer_delta = viewers_finale - viewers_premier,
    country_category = ifelse(country %in% list_top5_countries, country, "Other")
  ) %>%
  ggplot(aes(season, viewer_delta)) + 
  geom_line(aes(season, viewers_mean)) +
  ggrepel::geom_text_repel(
    data = filter(summary, season == 20),
    aes(season, viewers_mean), label = "Mean episode viewership\nper season"
    ) +
  geom_bar(aes(fill = country_category), stat = "identity", color = "gray20") +
  scale_fill_manual(values = countries_colors) +
  labs(
    title = "Difference between viewers at finale and viewers at premiere",
    x = "Season",
    y = "Viewers (millions)"
  ) + 
  theme_minimal() +
  theme(
    legend.position = "none",
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  )

# put the two plots together
plots_row <- plot_grid(sum_pie, sum_delta_plot, rel_widths = c(1,2))

title <- ggdraw() + 
  draw_label(
    "Survivor summary throughout the seasons",
    fontface = 'bold',
    x = 0,
    hjust = 0
  ) +
  theme(
    plot.margin = margin(0, 0, 0, 7)
  )

summary_plot <- plot_grid(
  title, plots_row,
  ncol = 1,
  rel_heights = c(0.1, 1)
)

summary_plot
```

Looks like overall mean viewership declined throughout the seasons (no surprise). Also, the variance in the delta between viewers at finale and viewers at premiere dropped off over time. I interpret this as demonstrating that the people who watched the later seasons of the show were already enfranchised, and were thus more likely to commit to finishing the season once started.

```{r}
repeat_winners <- filter(count(summary, vars = full_name), n > 1)$vars
seasons_with_returns <- nrow(filter(summary, str_detect(tribe_setup, "(returning|past)")))
```

Looking at the winners from each season, it is clear that there were only two repeat winners: `r repeat_winners`.

Of the `r nrow(summary)` seasons of Survivor, there were `r seasons_with_returns` seasons where returning players were allowed. From those `r seasons_with_returns` emerged `r length(repeat_winners)` repeat winners: `r repeat_winners`.

## Challenges

```{r}
knitr::kable(head(challenges),
             align = 'c',
             caption = "Head of challenges table")
```

Nothing here really interests me. Maybe could join this table with another to see relationships between `challenge type` and some other metrics (e.g. viewers, ratings, castaway personality type, etc). However, there are only two challenge types, both of which seem relatively evenly distributed. Specifically, there are `r nrow(filter(challenges, challenge_type == "reward"))` reward challenges and `r nrow(filter(challenges, challenge_type == "immunity"))` immnity challenges.

## Castaways

```{r}
knitr::kable(head(castaways),
             align = 'c',
             caption = "Head of castaways table")
```

### Age by Season

```{r}
castaways_sum <- castaways %>%
  group_by(season) %>%
  dplyr::summarise(
    players = n_distinct(full_name),
    mode_personality = Mode(personality_type),
    min_age = min(age),
    mean_age = mean(age),
    max_age = max(age)
  )
```

```{r}
castaways %>%
  ggplot(aes(season, age)) + 
  geom_boxplot(aes(group = season)) +
  labs(
    title = "Survivor: Distribution of castaway ages throughout the seasons",
    x = "Season",
    y = "Age"
  ) + 
  theme_minimal() + 
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  )
```

For the above boxplot, it is possible to assign color of the graph to season rating or viewers. However, this would be substantially confounded by the fact that both factors went down over time. At that point, it would be plotting two different mesaures (season by age, and season by viewers), and potentially suggesting a false causality. Therefore, this was not done.

### Result by Personality Type

```{r fig.height=8, fig.width=12}
personality_type_distr <- castaways %>%
  filter(!(is.na(personality_type))) %>%
  count(season, personality_type) %>%
  ggplot(aes(season, n, fill = personality_type)) + 
  geom_bar(stat = "identity", position = "fill") + 
  labs(
    title = "Personality type distribution\nby Survivor season",
    fill = "Legend",
    y = "",
    x = "Season"
  ) + 
  coord_flip() +
  scale_x_reverse() + 
  theme_minimal() + 
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_blank()
  )

castaway_result_noise <- filter(count(castaways, result), n < 10)$result

castaway_personality <- castaways %>%
  filter(
    !(result %in% castaway_result_noise),
    !(is.na(personality_type))
    ) %>%
  group_by(personality_type) %>%
  dplyr::summarise(
    `Percent of castaways who won the season` = 100*(sum(result == "Sole Survivor") / n()),
    `Percent of castaways who were finalists (winners and runner-ups)` = 100*(sum(str_detect(str_to_lower(result), "(runner-up|sole survivor)")) / n()),
    `Staying Power (sum of order squared, divided by count)` = sum(order^2) / n(),
  )

personality_success_plot <- castaway_personality %>%
  pivot_longer(cols = -personality_type, names_to = "success_metric", values_to = "rate") %>%
  ggplot(aes(reorder_within(personality_type, rate, success_metric), rate)) + 
  geom_bar(aes(fill = success_metric), stat = "identity") + 
  scale_x_reordered() + 
  facet_wrap(~ success_metric, ncol = 1, scales = "free") + 
  labs(
    title = "Which personality type was most succesful in Survivor?",
    x = "Personality Type",
    y = ""
  ) +
  theme_minimal() + 
  theme(
    legend.position = "none",
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  )

plot_grid(personality_success_plot, personality_type_distr, 
          rel_widths = c(2.7,1))
```

## Viewers

```{r}
knitr::kable(head(viewers),
             align = 'c',
             caption = "Viewers table head")
```

### Viewership and ratings over the course of a season

```{r fig.height=8, fig.width=12}
# viewers plot
viewers_season_plot <- viewers %>%
  filter(
    episode < 17,
    !(is.na(viewers))
  ) %>%
  group_by(episode) %>%
  dplyr::summarise(
    mean_viewers = mean(viewers),
    sd_viewers = sd(viewers)
  ) %>%
  ggplot(aes(episode, mean_viewers)) + 
  geom_line(color = "goldenrod4") + 
  geom_area(fill = "gold") + 
  coord_cartesian(ylim = c(12, 20)) + 
  labs(
    title = "Viewers throughout an average season",
    x = "Episode",
    y = "Average Viewers (millions)"
  ) + 
  theme_minimal()

# top viewed seasons
top_viewed_plots <- viewers %>%
  group_by(season_name, season) %>%
  dplyr::summarise(mean_viewers = mean(viewers, na.rm = TRUE)) %>%
  arrange(desc(mean_viewers)) %>%
  head(10) %>%
  mutate(
    season_tidy = paste0("S", season),
    season_name_tidy = str_remove(season_name, "Survivor: ")
  ) %>%
  ggplot(aes(fct_reorder(season_tidy, mean_viewers), mean_viewers)) + 
  geom_bar(stat = "identity", fill = "gold", color = "goldenrod4", alpha = 0.2) + 
  geom_text(aes(label = season_name_tidy, y = 0), size = 3, hjust = "left", color = "goldenrod4") + 
  geom_text(aes(label = round(mean_viewers)), hjust = 1.5, color = "goldenrod4") + 
  coord_flip() + 
  labs(
    title = "Most Viewed Season (millions)",
    x = "",
    y = ""
  ) + 
  theme_minimal() + 
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_blank()
  )

# rating plot
ratings_season_plot <- viewers %>%
  filter(
    episode < 17, 
    !(is.na(rating_18_49))
  ) %>%
  group_by(episode) %>%
  dplyr::summarise(mean_rating = mean(rating_18_49)) %>%
  ggplot(aes(episode, mean_rating)) + 
  geom_line() + 
  geom_area(fill = "darkorange", color = "darkorange4") + 
  coord_cartesian(ylim = c(3.4, 4.4)) + 
  labs(
    title = "Ratings throughout an average season",
    x = "Episode",
    y = "Average Rating"
  ) + 
  theme_minimal()

# top rated seasons
top_rated_seasons_plot <- viewers %>%
  group_by(season_name, season) %>%
  dplyr::summarise(mean_rating = mean(rating_18_49, na.rm = TRUE)) %>%
  arrange(desc(mean_rating)) %>%
  head(10) %>%
  mutate(
    season_tidy = paste0("S", season),
    season_name_tidy = str_remove(season_name, "Survivor: "),
    ) %>%
  ggplot(aes(fct_reorder(season_tidy, mean_rating), mean_rating)) + 
  geom_bar(stat = "identity", fill = "darkorange", color = "darkorange4", alpha = 0.2) + 
  geom_text(aes(label = season_name_tidy, y = 0), size = 3, hjust = "left", color = "darkorange4") +
  geom_text(aes(label = round(mean_rating, 1)), hjust = 1.5, color = "darkorange4") +
  coord_flip() + 
  labs(
    title = "Highest Rated Seasons",
    x = "",
    y = ""
  ) + 
  theme_minimal() + 
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_blank()
  )

# merge all four plots together
plots_row <- plot_grid(viewers_season_plot, top_viewed_plots, ratings_season_plot, top_rated_seasons_plot,
          ncol = 2, rel_widths = c(2.2,1))

title <- ggdraw() + 
  draw_label(
    "Survivor viewers and ratings by the season",
    fontface = "bold", x = 0, hjust = 0
  ) + 
  theme(
    plot.margin = margin(0,0,0,7)
  )

viewers_plot <- plot_grid(title, plots_row, ncol = 1, rel_heights = c(0.1,1))
viewers_plot
```
