---
title: "Bees"
author: "Nick Cruickshank"
date: "1/15/2022"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This week's [Tidy Tuesday Project](https://github.com/rfordatascience/tidytuesday/tree/master/data/2022/2022-01-11) features data concerning bee colonies, kudos to [G. Karamanis](https://github.com/rfordatascience/tidytuesday/issues/404).

> This report provides information on honey bee colonies in terms of number of colonies, maximum, lost, percent lost, added, renovated, and percent renovated, as well as colonies lost with Colony Collapse Disorder symptoms with both over and less than five colonies. The report also identifies colony health stressors with five or more colonies. The data for operations with honey bee colonies are collected from a stratified sample of operations that responded as having honey bees on the Bee and Honey Inquiry and from the NASS list frame. For operations with five or more colonies, data was collected on a quarterly basis; operations with less than five colonies were collected with an annual survey.

# Analysis

```{r}
# libraries
library(broom) # for fortify
library(here)
library(ggbump) # for geom_sigmoid()
library(ggtext)
library(glue)
library(geojsonio) # for geosjaon_read()
library(readr)
library(tidyverse)
```

```{r}
# data
here <- here()

colony <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-01-11/colony.csv')
stressor <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-01-11/stressor.csv')
spdf <- geojson_read(glue("{here}/2022/2022-01-11 - Bees/us_states_hexgrid.geojson"), what = "sp")

spdf@data = spdf@data %>% mutate(google_name = gsub(" \\(United States\\)", "", google_name))
spdf_fortified <- tidy(spdf, region = "google_name")
```

```{r}
head(colony)
```

```{r}
head(stressor)
```

# Ideas

## Timeline of Stressors

Timeline of stressors over time across all states. Maybe roll states up into US regions.

## Hexbin Map of US

Maybe show changes in stressor prevalence between 2015 and 2021. Sigmoid to reference additional metrics on the right.

Here is a [Good Article](https://www.r-graph-gallery.com/328-hexbin-map-of-the-usa.html) describing how to make hexbins in R.

## EDA

```{r}
count(colony, months)
```

```{r}
paste0("Min Year: ", min(colony$year), ", Max Year: ", max(colony$year))
```

## Hex  Bins

### Tidy

```{r}
# values
sigmoid_step_size <- 10
```

```{r}
# transform spdf `id` into state abb
## create dataframe for mapping
states <- tibble(
  id = append(state.name, "District of Columbia"),
  abb = append(state.abb, "DC")
)

## join together
spdf_fort2 <- spdf_fortified %>%
  select(id, long, lat) %>%
  left_join(states, by = "id")

spdf_centers <- spdf_fort2 %>%
  group_by(abb) %>%
  dplyr::summarise(
    # find centers of hexes for state names
    mid_long = mean(long),
    mid_lat = mean(lat),
    # and slightly above center to start the sigmoid
    sig_long = mid_long,
    sig_lat = mid_lat + 1
  )

## define sigmoid

# define plotting metrics for state hexbins
state_colony_growth <- colony %>%
  filter(state != "United States") %>%
  mutate(
    colony_added = replace_na(colony_added, 0),
    colony_lost = replace_na(colony_lost, 0),
    growth = colony_added - colony_lost
  ) %>%
  group_by(state) %>%
  dplyr::summarise(total_growth = sum(growth, na.rm = TRUE)) %>%
  ungroup() %>%
  rename(c("id" = "state")) %>%
  left_join(states, by = "id") %>%
  mutate(
    abs_growth = abs(total_growth),
    growth_bin = case_when(
      total_growth < -200000 ~ "Major Decline",
      total_growth < -5000 ~ "Decline",
      total_growth < 5000 ~ "Steady",
      total_growth < 200000 ~ "Growth",
      total_growth >= 20000 ~ "Major Growth"
    ),
    growth_bin = factor(growth_bin, levels = c("Major Decline", "Decline", "Steady", "Growth", "Major Growth"))
  )

# join states and spdf_fort2 for plotting purposes
spdf_bees <- spdf_fort2 %>%
  left_join(state_colony_growth, by = c("id", "abb"))

# highlight the greatest movers for sigmoid plotting
highlight_states <- state_colony_growth %>%
  filter(abs_growth >= 100000) %>%
  left_join(spdf_centers, by = "abb") %>%
  arrange(growth_bin) %>%
  mutate(
    sig_xend = -50,
    sig_yend = rev(seq(25, 75, sigmoid_step_size)), # rev because that incidentally lines up with long better
    text_color = case_when(
      growth_bin == "Major Decline" ~ "darkgoldenrod4",
      growth_bin == "Decline" ~ "darkgoldenrod2",
      growth_bin == "Steady" ~ "gold",
      growth_bin == "Growth" ~ "greenyellow",
      growth_bin == "Major Growth" ~ "yellowgreen"
    )
  )
```

```{r}
# define sigmoid plots of stressors over time

## define x points for each quarter
stressor_x <- stressor %>%
  rename(c("id" = "state")) %>%
  inner_join(states, by = "id") %>%
  mutate(
    quarter = case_when(
      months == "January-March" ~ paste0(year, "-Q1"),
      months == "April-June" ~ paste0(year, "-Q2"),
      months == "July-September" ~ paste0(year, "-Q3"),
      months == "October-December" ~ paste0(year, "-Q4")
    )
  ) %>%
  #arrange(abb, quarter) %>%
  count(abb, quarter) %>%
  group_by(abb) %>%
  mutate(x = seq(1,52,2)) %>%
  select(-n)

## define the max stress value for scaling purposes
max_stress <- stressor %>%
  filter(state %in% highlight_states$id) %>%
  arrange(desc(stress_pct)) %>%
  head(1) %>%
  pull(stress_pct)

## final tidying for plotting purposes
sig_stressor <- stressor %>%
  rename(c("id" = "state")) %>%
  inner_join(highlight_states, by = "id") %>%
  mutate(
    quarter = case_when(
      months == "January-March" ~ paste0(year, "-Q1"),
      months == "April-June" ~ paste0(year, "-Q2"),
      months == "July-September" ~ paste0(year, "-Q3"),
      months == "October-December" ~ paste0(year, "-Q4")
    )
  ) %>%
  left_join(stressor_x, by = c("abb", "quarter")) %>%
  mutate(
    x_adj = x + sig_xend + 20,
    stress_pct_scaled = (stress_pct / 100) * (sigmoid_step_size - 3),
    y = stress_pct_scaled + sig_yend
  ) %>%
  group_by(abb, quarter) %>%
  filter(stress_pct == max(stress_pct)) %>%
  ungroup() %>%
  mutate(
    stressor_color = case_when(
      stressor == "Other pests/parasites" ~ "grey40",
      stressor == "Pesticides" ~ "cyan",
      stressor == "Unknown" ~ "grey60",
      stressor == "Varroa mites" ~ "magenta"
    )
  )

## define x axis label coordinates
x_lab_coords <- sig_stressor %>%
  distinct(year, quarter, x_adj, sig_yend) %>%
  group_by(year, sig_yend) %>%
  dplyr::summarise(x_lab = mean(x_adj)) %>%
  ungroup() %>%
  mutate(ylab = sig_yend - 1) %>%
  filter(year %% 2 == 0)

## define ax lines for plot
sig_ax_lines <- sig_stressor %>%
  distinct(abb, sig_xend, sig_yend) %>%
  mutate(
    xstart = sig_xend + 20,
    xend = 20, # xstart + 102,
    ystart = sig_yend,
    yend = ystart + (sigmoid_step_size - 3)
  ) %>%
  select(-sig_xend, -sig_yend)
```

### Plot

```{r Bee Growth and Stress, fig.height=9, fig.width=13}
ggplot() +
  annotate("text", x = -125, y = 95, label = "Growth & Stress in Bee Colonies", color = "grey90", size = 15, fontface = "bold", hjust = 0) +
  annotate("text", x = -125, y = 88, label = "Growth = Colonies Added - Colonies Lost", color = "grey90", size = 8, hjust = 0) +
  annotate("text", x = -125, y = 83, label = "Stress = Percent of colony affected by stressor", color = "grey90", size = 8, hjust = 0) +
  annotate("text", x = -115, y = 78, label = "Most Common Stressors:", color = "grey90", size = 6, hjust = 0) + 
  annotate("text", x = -105, y = 74, label = "Pesticides", color = "cyan", size = 6, fontface = "italic", hjust = 0) + 
  annotate("text", x = -105, y = 71, label = "Varroa mites", color = "magenta", size = 6, fontface = "italic", hjust = 0) + 
  
  # state polygons
  annotate("text", x = -100, y = 54, label = "Total Growth\n2015 to 2021", color = "grey90", size = 5, fontface = "bold") +
  
  ## plot polygons
  geom_polygon(data = spdf_bees, aes(x = long, y = lat, fill = growth_bin, group = abb), color="grey10") +
  geom_point(data = spdf_centers, aes(x = sig_long, y = sig_lat, group = abb), color = "grey10") + 
  geom_text(data=spdf_centers, aes(x=mid_long, y=mid_lat, label=abb), color = "grey10") + 
  scale_fill_manual(values = c(
    "Major Decline" = "darkgoldenrod4",
    "Decline" = "darkgoldenrod2",
    "Steady" = "gold",
    "Growth" = "greenyellow",
    "Major Growth" = "yellowgreen"
  ), guide = "none") +
  
  # sigmoid to timelines details
  geom_sigmoid(data = highlight_states, aes(x = sig_long, xend = sig_xend, y = sig_lat, yend = sig_yend, group = abb), color = "grey90") +
  geom_richtext(
    data = highlight_states, 
    aes(x = sig_xend + 2, y = sig_yend, label = paste0("<b>", abb, "</b><br><b>", growth_bin, "</b><br>", round(total_growth / 1000), "k Colonies"), color = text_color),
    label.color = NA, fill = NA, label.padding = grid::unit(rep(0,4), "pt"), hjust = 0, size = 3, vjust = 0
  ) +
  
  # sigmoid time plots
  annotate("text", x = mean(x_lab_coords$x_lab), y = 85, label = "Greatest Quarterly\nStressors", color = "grey90", size = 5, fontface = "bold") + 
  
  ## sigmoid axis
  
  ### x axis
  geom_linerange(data = sig_ax_lines, aes(xmin = xstart, xmax = xend, y = ystart), color = "grey90") + 
  geom_text(data = x_lab_coords, aes(x = x_lab, y = ylab, label = year), color = "grey90", size = 2.5) +
  
  ### y axis
  geom_linerange(data = sig_ax_lines, aes(x = xstart, ymin = ystart, ymax = yend), color = "grey90") + 
  geom_text(data = sig_ax_lines, aes(x = xstart - 3, y = ystart, label = "0%"), color = "grey90", size = 2.5) + 
  geom_text(data = sig_ax_lines, aes(x = xstart - 3, y = (ystart + yend) / 2, label = "50%"), color = "grey90", size = 2.5) + 
  geom_text(data = sig_ax_lines, aes(x = xstart - 3, y = yend, label = "100%"), color = "grey90", size = 2.5) + 
  
  ## siqmoid points
  geom_line(data = sig_stressor, aes(x_adj, y, group = abb), color = "grey80") +
  geom_point(data = sig_stressor, aes(x_adj, y, color = stressor_color)) +
  
  scale_color_identity() +
  # general scales, themes, etc
  coord_cartesian(
    xlim = c(min(spdf_fort2$long) - 2, 25), 
    ylim = c(min(spdf_fort2$lat) - 5, 100),
    expand = c(0,0)
  ) +
  labs(caption = "Data Source: <b>USDA</b> |  Visualization: <b>N. Cruickshank</b> | #TidyTuesday") +
  theme_void() + 
  theme(
    plot.background = element_rect(fill = "grey10"),
    plot.caption = element_textbox(hjust = 0.5, size = 8, color = "grey90")
  )
```
