---
title: "Cricket"
author: "Nick Cruickshank"
date: "12/4/2021"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

# Introduction

This week's [Tidy Tuesday Project](https://github.com/rfordatascience/tidytuesday/tree/master/data/2021/2021-11-30) features data about [Cricket](https://en.wikipedia.org/wiki/Cricket) from [ESPN Cricket](https://www.espncricinfo.com/).

> Cricket is a bat-and-ball game played between two teams of eleven players on a field at the centre of which is a 22-yard (20-metre) pitch with a wicket at each end, each comprising two bails balanced on three stumps. The game proceeds when a player on the fielding team, called the bowler, "bowls" (propels) the ball from one end of the pitch towards the wicket at the other end. The batting side's players score runs by striking the bowled ball with a bat and running between the wickets, while the bowling side tries to prevent this by keeping the ball within the field and getting it to either wicket, and dismiss each batter (so they are "out"). Means of dismissal include being bowled, when the ball hits the stumps and dislodges the bails, and by the fielding side either catching a hit ball before it touches the ground, or hitting a wicket with the ball before a batter can cross the crease line in front of the wicket to complete a run. When ten batters have been dismissed, the innings ends and the teams swap roles. The game is adjudicated by two umpires, aided by a third umpire and match referee in international matches.

> A One Day International (ODI) is a form of limited overs cricket, played between two teams with international status, in which each team faces a fixed number of overs, currently 50, with the game lasting up to 9 hours.[1][2] The Cricket World Cup, generally held every four years, is played in this format. One Day International matches are also called Limited Overs Internationals (LOI), although this generic term may also refer to Twenty20 International matches. They are major matches and considered the highest standard of List A, limited-overs competition.

> The Cricket World Cup (officially known as ICC Men's Cricket World Cup)[2] is the international championship of One Day International (ODI) cricket. The event is organised by the sport's governing body, the International Cricket Council (ICC), every four years, with preliminary qualification rounds leading up to a finals tournament. The tournament is one of the world's most viewed sporting events and is considered the "flagship event of the international cricket calendar" by the ICC

# Analysis

```{r libraries}
# libraries
library(forcats)
library(ggbump)
library(ggnetwork)
library(glue)
library(readr)
library(tidyverse)
```

```{r data}
# data
matches <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-11-30/matches.csv')
```

```{r}
# head
head(matches)
```

## EDA

### Most Represented Teams

Pretty big cut off after the top 9 teams.

```{r}
team_counts <- matches %>%
  select(team1, team2) %>%
  pivot_longer(cols = c("team1", "team2")) %>%
  count(value, sort = TRUE)

relevant_teams <- team_counts %>%
  head(10) %>%
  pull(value)

team_counts %>%
  ggplot(aes(n, fct_reorder(value, n))) + 
  geom_bar(stat = "identity") + 
  labs(title = "Very big cutoff after top 11 teams")
```

### Distribution of Scores

```{r}
score_dist <- matches %>%
  select(score_team1, score_team2) %>%
  pivot_longer(cols = c("score_team1", "score_team2")) 

shapiro.test(score_dist$value)

score_dist %>%
  ggplot() + 
  geom_histogram(aes(value)) + 
  labs(
    title = "Distribution of Scores",
    subtitle = "Normally Distributed (p-value = 0.0001127)"
  )
```

### Which teams have the highest win rate?

```{r}
matches %>%
  select(team1, team2, winner) %>%
  pivot_longer(cols = c("team1", "team2")) %>%
  mutate(win = ifelse(winner == value, 1, 0)) %>%
  group_by(value) %>%
  dplyr::summarise(
    games = n(),
    wins = sum(win, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  mutate(
    win_rate = wins / games,
    disp = ifelse(win_rate > 0.5, "Positive", "Negative")
  ) %>%
  filter(win_rate > 0) %>%
  arrange(desc(win_rate)) %>%
  ggplot(aes(win_rate, fct_reorder(value, win_rate))) + 
  geom_vline(xintercept = 0.5, linetype = "dashed") +
  geom_bar(aes(color = disp), fill = "white", stat = "identity", size = 2) + 
  geom_text(aes(label = paste(round(100 * win_rate, 2), "% (", wins, " out of ", games, " games)")), 
            hjust = -0.1, color = "grey30") + 
  scale_color_manual(values = c(
    "Negative" = "red",
    "Positive" = "green3"
  )) + 
  scale_y_discrete(name = NULL) + 
  scale_x_continuous(limits = c(0, 1.25), expand = c(0,0), name = NULL) + 
  labs(
    title = "Only 5 of the 21 represented teams have a positive win rate"
  ) +
  theme_minimal() + 
  theme(
    legend.position = "none",
    panel.grid = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.x = element_blank()
  )
```

### Is the player of the match more often on the winnning team, or the losing team?

Perhaps not surprisingly, the Player of the Match is most often on the winning team.

```{r}
matches %>%
  select(winner, player_of_match_team) %>%
  mutate(disp = ifelse(winner == player_of_match_team, "Winner", "Loser")) %>%
  count(disp, sort = TRUE)
``` 

### When and where do most games happen?

```{r}
matches %>%
  count(time_of_day, sort = TRUE)
```

```{r}
matches %>%
  count(venue, sort = TRUE)
```

## Tidy

```{r tidy}
# tidy

# # create team 1 df
# team1 <- matches %>%
#   select(match_id, team1, score_team1, wickets_team, team1_away_or_home) %>%
#   rename(c(
#     "team" = "team1",
#     "score" = "score_team1",
#     "wickets" = "wickets_team",
#     "away_or_home" = "team1_away_or_home"
#   )) %>%
#   mutate(team_number = "team1")
# 
# # create team 2 df
# team2 <- matches %>%
#   select(match_id, team2, score_team2, wickets_team2, team2_home_away) %>%
#   rename(c(
#     "team" = "team2",
#     "score" = "score_team2",
#     "wickets" = "wickets_team2",
#     "away_or_home" = "team2_home_away"
#   )) %>%
#   mutate(team_number = "team2")
# 
# # create match df
# matches2 <- matches %>%
#   select(match_id, winner, margin, series, player_of_match, player_of_match_team, 
#          venue, toss, toss_decision, ground, ground_city, ground_country)
# 
# # join all together
# bind_rows(team1, team2) %>%
#   left_join(matches2, by = "match_id") %>%
#   filter(
#     team %in% relevant_teams
#   ) %>%
#   group_by(match_id, winner, margin, series, player_of_match, player_of_match_team,
#          venue, toss, toss_decision, ground, ground_city, ground_country) %>%
#   summarize(pair = glue_collapse(sort(team), sep = "-")) %>%
#   ungroup() %>%
#   count(pair)


match_pairs <- matches %>%
  select(match_id, team1, team2) %>%
  pivot_longer(cols = c("team1", "team2"), names_to = "team_number", values_to = "team") %>%
  filter(team %in% relevant_teams) %>%
  group_by(match_id) %>%
  summarize(pair = as.character(glue_collapse(sort(team), sep = "-"))) %>%
  ungroup() %>%
  mutate(dash = str_count(pair, "-")) %>%
  filter(dash > 0) %>%
  select(-dash)

# plot coordinates by team
team_coords <- tibble(
  team1 = sort(relevant_teams),
  team2 = sort(relevant_teams),
  coord = c("2,1", "4,1", "5,3", "5,5", "5,7", "4,9", "2,9", "1,7", "1,5", "1,3")
) 

network <- matches %>%
  inner_join(match_pairs, by = "match_id") %>%
  select(pair, margin, winner) %>%
  separate(col = "pair", into = c("team1", "team2"), sep = "-", remove = FALSE) %>%
  mutate(side = ifelse(winner == team1, "left", "right")) %>%
  pivot_longer(cols = c("team1", "team2")) %>%
  group_by(pair, side) %>%
  dplyr::summarise(
    margin_adj = mean(margin, na.rm = TRUE)
  ) %>% 
  ungroup() %>%
  pivot_wider(id_cols = "pair", values_from = margin_adj, names_from = "side") %>%
  separate(pair, into = c("team1", "team2"), sep = "-", remove = FALSE) %>%
  mutate(
    left = replace_na(left, 0),
    right = replace_na(right, 0),
    margin_delta = abs(left - right),
    in_favor_of = ifelse(left - right > 0, team1, team2)
  ) %>%
  select(-left, -right) %>%
  left_join(select(team_coords, team1, coord), by = "team1") %>%
  rename(c("team1_coord" = "coord")) %>%
  left_join(select(team_coords, team2, coord), by = "team2") %>%
  rename(c("team2_coord" = "coord")) %>%
  separate(team1_coord, into = c("x", "y"), sep = ",") %>%
  separate(team2_coord, into = c("xend", "yend"), sep = ",") %>%
  mutate(
    x = as.double(x),
    y = as.double(y),
    xend = as.double(xend),
    yend = as.double(yend),
    margin_bin = case_when(
      margin_delta < 20 ~ "0 - 20",
      margin_delta < 40 ~ "20 - 40",
      margin_delta >= 40 ~ "40+"
    )
  )

team_coords2 <- team_coords %>%
  select(team1, coord) %>%
  separate(coord, into = c("x", "y"), sep = ",") %>%
  mutate(x = as.double(x), y = as.double(y))
```

Delta = the difference in the average margin when the left team wins as opposed to the average margin when the right team wins.
 - High Positive Value = Left wins by a wider margin when they win
 - High Negative Value = Right wins by a wider margin when they win

## Visualization

Create a network plot where each node is a team and the edges are the average margin between those teams

```{r}
# network plot of teams
network %>%
  ggplot() +
  geom_edges(aes(x = x, y = y, xend = xend, yend = yend, color = in_favor_of, size = margin_bin, linetype = margin_bin), curvature = 0.2) +
  #geom_nodes(aes(x, y, color = team1), size = 10) + 
  geom_point(data = team_coords2, aes(x, y, color =  team1), size = 10) +
  ggrepel::geom_text_repel(
    data = team_coords2, 
    aes(x, y, label = team1),
    hjust = 0, color = "white"
  ) + 
  coord_cartesian(xlim = c(0, 6), ylim = c(0, 10), expand = c(0,0)) + 
  scale_color_brewer(palette = "Paired", guide = "none") + 
  scale_size_manual(values = c(
    "0 - 20" = 1,
    "20 - 40" = 2,
    "40+" = 3
  )) +
  scale_linetype_manual(values = c(
    "0 - 20" = "dashed",
    "20 - 40" = "longdash",
    "40+" = "solid"
  )) +
  theme_void() + 
  theme(
    legend.position = "top",
    legend.text = element_text(color = "white"),
    plot.background = element_rect(fill = "grey20")
  )
```

