---
title: "20210831 - Bird Baths"
author: "Nick Cruickshank"
date: "9/2/2021"
output: 
  github_document:
    toc: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

# Introduction

# Exploratory Data Analysis

```{r libraries}
# libraries
library(forcats)
library(readr)
library(tidyverse)
```

```{r data}
# data
bird_baths <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-08-31/bird_baths.csv')

## view dataset
bird_baths %>%
  head()
```

## What are the most populous birds?

```{r}
bird_counts <- bird_baths %>%
  group_by(bird_type) %>%
  summarise(total = sum(bird_count)) %>%
  arrange(desc(total)) 

# bird_counts %>%
#   ggplot(aes(fct_reorder(bird_type, total), total)) + 
#   geom_hline(yintercept = 300) + 
#   geom_point() + 
#   scale_x_discrete(name = NULL, labels = NULL)

most_populous_birds <- bird_counts %>%
  filter(total > 300) %>%
  pull(bird_type)
```

## Which birds had the biggest raw change from Urban to Rural?

Looks like there's a fairly decent breaking point around 300 birds total.

```{r}

```

## Which birds had the biggest raw change from 2014 to 2015?

For either urban and / or rural.

## Ideas from Cleary et al 2016

### Do assemblages at urban and rural bird baths differ across bioregions?

Use [PERMANOVA](https://en.wikipedia.org/wiki/Permutational_analysis_of_variance), which can be found in the `vegan` and `InPerm` packages.


### Do assemblages at urban and rural bird baths differ within bioregions?

[Multidimensional Scaling (nMDS) ordination techniques](http://strata.uga.edu/8370/lecturenotes/multidimensionalScaling.html) found in library `vegan` and PERMANOVA

### Is species richness different at bird baths in rural areas compared to urban areas?

