---
title: "20210831 - Bird Baths"
author: "Nick Cruickshank"
date: "9/2/2021"
output: 
  github_document:
    toc: true
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

# Introduction

# Exploratory Data Analysis

```{r libraries}
# libraries
library(forcats)
library(ggrepel)
library(ggtext)
library(patchwork)
library(readr)
library(tidyverse)
```

```{r data}
# data
bird_baths <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-08-31/bird_baths.csv')

## view dataset
bird_baths %>%
  head()
```

## What are the most populous birds?

```{r}
bird_counts <- bird_baths %>%
  group_by(bird_type) %>%
  summarise(total = sum(bird_count)) %>%
  arrange(desc(total)) 

# bird_counts %>%
#   ggplot(aes(fct_reorder(bird_type, total), total)) + 
#   geom_hline(yintercept = 300) + 
#   geom_point() + 
#   scale_x_discrete(name = NULL, labels = NULL)

most_populous_birds <- bird_counts %>%
  filter(total > 300) %>%
  pull(bird_type)

most_populous_birds
```

## Which birds had the biggest raw change from Urban to Rural?

Looks like there's a fairly decent breaking point around 300 birds total.



## Which birds had the biggest raw change from 2014 to 2015?

For either urban and / or rural.

Overall change (ignoring urban / rural and bioregion) shows the following reductions from 2014

```{r}
# get list of most frequently occuring birds
top_10_birds <- bird_baths %>%
  filter(!(is.na(survey_year))) %>%
  group_by(bird_type) %>%
  dplyr::summarise(total = sum(bird_count)) %>%
  ungroup() %>%
  arrange(desc(total)) %>%
  head(10) %>%
  pull(bird_type)

# may want to consider using the species with the most statistically significant changes from one group to other
```

```{r}
# create seqmnet df
bath_years <- bird_baths %>%
  filter(
    !(is.na(survey_year))
  ) %>%
  group_by(survey_year, urban_rural, bird_type) %>%
  dplyr::summarise(
    total_birds = sum(bird_count, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  pivot_wider(id_cols = c("urban_rural", "bird_type"), names_from = "survey_year", values_from = "total_birds", names_prefix = "y") %>%
  group_by(urban_rural, bird_type) %>%
  mutate(
    color = ifelse(y2014 > y2015, "#876429", "#298735"),
    bird_id = cur_group_id()
  ) %>%
  ungroup()
```

```{r Rural Bird Bath Lollipops, fig.height=10, fig.width=5.625}
# plot
rural_lollipops <- bath_years %>%
  filter(urban_rural == "Rural", bird_type %in% top_10_birds) %>% 
  ggplot() +
  # total count axis
  annotate("segment", x = 1.1, xend = 1.9, y = seq(10, 70, 5), yend = seq(10, 70, 5), alpha = 0.2, size = 0.2) +
  annotate("point", x = 1.5, y = seq(10, 70, 10), size = 16, color = "grey96") +
  annotate("text", x = 1.5, y = seq(10, 70, 10), label = seq(10, 70, 10), size = 8, alpha = 0.2, fontface = "bold", color = "#87297B") +
  # year annotations
  annotate("text", x = c(0.9, 2.1), y = 5, label = c("2014", "2015"), hjust = c(1,0), fontface = "bold", size = 10, alpha = 0.3, color = "#87297B") + 
  # lollipops
  geom_segment(aes(y = y2014, yend = y2015, x = 1, xend = 2, color = color)) + 
  geom_point(aes(y = y2014, x = 1, color = color, size = y2014)) + 
  geom_point(aes(y = y2015, x = 2, color = color, size = y2015)) + 
  # bird labels
  ggrepel::geom_text_repel(
    data = filter(bath_years, urban_rural == "Rural", bird_type %in% top_10_birds, bird_id %% 2 == 0),
    aes(y = y2014, x = 1, label = bird_type, color = color),
    nudge_x = -0.2, hjust = 1, direction = "y", point.padding = 1, segement.size = 0.2, fontface = "italic"
  ) +
  ggrepel::geom_text_repel(
    data = filter(bath_years, urban_rural == "Rural", bird_type %in% top_10_birds, bird_id %% 2 != 0),
    aes(y = y2015, x = 2, label = bird_type, color = color),
    nudge_x = 0.2, hjust = 0, direction = "y", point.padding = 1, segement.size = 0.2, fontface = "italic"
  ) +
  # scales, themes, etc
  scale_color_identity() + 
  scale_size_continuous(guide = "none") + 
  scale_x_continuous(limits = c(0.5, 2.5)) + 
  labs(
    title = "RURAL"
  ) + 
  theme_void() + 
  theme(
    plot.title = element_textbox(color = "#294C87", hjust = 0.5, size = 24, face = "bold")
  )

rural_lollipops
```

```{r Urban Bird Bath Lollipops, fig.height=10, fig.width=5.625}
# plot
urban_lollipops <- bath_years %>%
  filter(urban_rural == "Urban", bird_type %in% top_10_birds) %>% 
  ggplot() +
  # total count axis
  annotate("segment", x = 1.1, xend = 1.9, y = seq(10, 140, 10), yend = seq(10, 140, 10), alpha = 0.2, size = 0.2) +
  annotate("point", x = 1.5, y = seq(10, 140, 10), size = 16, color = "grey96") +
  annotate("text", x = 1.5, y = seq(10, 140, 10), label = seq(10, 140, 10), size = 5, alpha = 0.2, fontface = "bold", color = "#87297B") +
  # year annotations
  annotate("text", x = c(0.9, 2.1), y = 5, label = c("2014", "2015"), hjust = c(1,0), fontface = "bold", size = 10, alpha = 0.3, color = "#87297B") + 
  # lollipops
  geom_segment(aes(y = y2014, yend = y2015, x = 1, xend = 2, color = color)) + 
  geom_point(aes(y = y2014, x = 1, color = color, size = y2014)) + 
  geom_point(aes(y = y2015, x = 2, color = color, size = y2015)) + 
  # bird labels
  ggrepel::geom_text_repel(
    data = filter(bath_years, urban_rural == "Urban", bird_type %in% top_10_birds, bird_id %% 2 == 0),
    aes(y = y2014, x = 1, label = bird_type, color = color),
    nudge_x = -0.2, hjust = 1, direction = "y", point.padding = 1, segement.size = 0.2, fontface = "italic"
  ) +
  ggrepel::geom_text_repel(
    data = filter(bath_years, urban_rural == "Urban", bird_type %in% top_10_birds, bird_id %% 2 != 0),
    aes(y = y2015, x = 2, label = bird_type, color = color),
    nudge_x = 0.2, hjust = 0, direction = "y", point.padding = 1, segement.size = 0.2, fontface = "italic"
  ) +
  # scales, themes, etc
  scale_color_identity() + 
  scale_size_continuous(guide = "none") + 
  scale_x_continuous(limits = c(0.5, 2.5)) + 
  labs(
    title = "URBAN"
  ) + 
  theme_void() + 
  theme(
    plot.title = element_textbox(color = "#294C87", hjust = 0.5, size = 24, face = "bold")
  )

urban_lollipops
```

```{r Bird Lollipops, fig.height=10, fig.width=9}
# both plots, side by side
## doesn't look amazing so far
# urban_lollipops + rural_lollipops + 
#   plot_layout(ncol = 2)
```


## Ideas from Cleary et al 2016

### Do assemblages at urban and rural bird baths differ across bioregions?

Use [PERMANOVA](https://en.wikipedia.org/wiki/Permutational_analysis_of_variance), which can be found in the `vegan` and `InPerm` packages.

### Do assemblages at urban and rural bird baths differ within bioregions?

[Multidimensional Scaling (nMDS) ordination techniques](http://strata.uga.edu/8370/lecturenotes/multidimensionalScaling.html) found in library `vegan` and PERMANOVA

### Is species richness different at bird baths in rural areas compared to urban areas?

